Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2019-01-26 paper/paper.tex
Fit onto 8 pages again RC
Revise paper according to Reviews
- beware of the comma after the superscript numbers in the author list:
  you can remove it, just check the LaTeX template -> this is
  intentionally
- in the abstract, please prevent the line break between Tier and 1 (done)
- a reference to the GeliX Nebula Science Cloud project might be nice
(added https://www.hnscicloud.eu/ as footnote)
- "eclipse plugin": what is eclipse (and shoudn't it be capitalized)? ->
  Ecplise IDE
- "The MONARC project was started 1998" -> "in 1998"

	This paper presents a proof of concept in which the computing jobs at the Tier 1 center GridKa are modeled and simulated. To model the computing jobs we extended the Palladio simulator with a mechanism to simulate load balancing strategies. Furthermore, we implemented an automated model parameter analysis and model creation.

@@

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources for user jobs. Cloud resources, such as the Helix Nebula Science Cloud and Amazon Web Services, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs. This results in a more complex scheduling, because each resource has different optimal job types and mixes. Furthermore, a more dynamic data placement, where fewer centers host data, leads to an increase in the number of remote data accesses.

@@

\section{Foundations: Palladio}
\label{sec:palladio}
Palladio is a model driven software architecture simulator, which is implemented as an eclipse plugin and developed by Karlsruhe Institute of Technology (KIT), FZI Research Center for Information Technology, and University of Paderborn. Palladio predicts quality of software properties (e.g. performance) using several models of a system \cite{BECKER20093}. These models are: 
\begin{itemize}
	\item The component model: Specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
	\item The assembly model: Represents how the components are connected to model the software architecture.
	\item The resource model: Describes the resource environment. It contains the number and characteristics of the resource containers on which the components could be deployed and the topology of the network connecting these containers.
	\item The allocation model: Represents the mapping of the different components to the resource containers.
	\item The usage model: Defines the usage of a system regarding workload, user behavior, and parameters.

\end{itemize}
Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.

@@

\begin{figure}[h]
	\centering
	\vspace{-10px}
	\begin{subfigure}{0.48\linewidth}
		\includegraphics[width=\linewidth]{images/validation/throughputs}
	\end{subfigure}
	\begin{subfigure}{0.48\linewidth}
		\includegraphics[width=\linewidth]{images/validation/utilizations}
	\end{subfigure}
	\vspace{-10px}
	\caption{Measured and simulated throughputs and CPU efficiency using monitoring data from May 2018}
	\label{fig:validation}
\end{figure}
\vspace{10px}
\noindent To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types, and the overall CPU efficiency of the site.
These metrics describe how well the simulation matches the measured performance data.
Comparisons between simulation results and measurements from the system are shown in \Cref{fig:validation}.
One result is that simulated throughput matches the reference with a difference of about 5.6\%.
Furthermore, the shares of the simulated and measured job types also match closely.

@@

\section{Related Work}
\label{related}
The MONARC project was started 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
However, after the initial design was developed, the project was discontinued in 2000 and thus no model and simulation of the WLCG was created. 
In 2004 a second version of the MONARC simulator was released with several improvements, such as simulation of data replication. However, MONARC2 showed scalability issues when simulating a Tier-2 site and accuracy issues regarding network load \cite{1742-6596-331-7-072038}.

>>>>

	This paper presents a proof of concept in which the computing jobs at the \mbox{Tier 1} center GridKa are modeled and simulated. To model the computing jobs we extended the Palladio simulator with a mechanism to simulate load balancing strategies. Furthermore, we implemented an automated model parameter analysis and model creation.

@@

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources for user jobs. Cloud resources, such as the Helix Nebula Science Cloud\footurl{http://www.helix-nebula.eu/} and Amazon Web Services, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs. This results in a more complex scheduling, because each resource has different optimal job types and mixes. Furthermore, a more dynamic data placement, where fewer centers host data, leads to an increase in the number of remote data accesses.

@@

\section{Foundations: Palladio}
\label{sec:palladio}
Palladio is a model driven software architecture simulator, which is implemented as a plugin for the Eclipse IDE and developed by Karlsruhe Institute of Technology (KIT), FZI Research Center for Information Technology, and University of Paderborn. Palladio predicts quality of software properties (e.g. performance) using the following models of a system \cite{BECKER20093}: 
\begin{itemize}
	\vspace{-0.2pc}
	\itemsep-0.2em
	\item The component model: Specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
	\item The assembly model: Represents how the components are connected to model the software architecture.
	\item The resource model: Describes the resource environment. It contains the number and characteristics of the resource containers on which the components could be deployed and the topology of the network connecting these containers.
	\item The allocation model: Represents the mapping of the different components to the resource containers.
	\item The usage model: Defines the usage of a system regarding workload, user behavior, and parameters.
\end{itemize}
	\vspace{-0.2pc}
Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.

@@

\begin{figure}[h]
	\centering
	\vspace{-10px}
	\begin{subfigure}{0.48\linewidth}
		\includegraphics[width=\linewidth]{images/validation/throughputs}
	\end{subfigure}
	\begin{subfigure}{0.48\linewidth}
		\includegraphics[width=\linewidth]{images/validation/utilizations}
	\end{subfigure}
	\vspace{-10px}
	\caption{Measured and simulated throughputs and CPU efficiency job data from May 2018}
	\label{fig:validation}
\end{figure}
\vspace{10px}
\noindent To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types, and the overall CPU efficiency of the site.
These metrics describe how well the simulation matches the measured performance data.
Comparisons between simulation results and measurements from the system are shown in \Cref{fig:validation}.
One result is that simulated throughput matches the reference with a difference of about 5.6\%.
Furthermore, the shares of the simulated and measured job types also match closely.

@@

\section{Related Work}
\label{related}
The MONARC project was started in 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
However, after the initial design was developed, the project was discontinued in 2000 and thus no model and simulation of the WLCG was created. 
In 2004 a second version of the MONARC simulator was released with several improvements, such as simulation of data replication. However, MONARC2 showed scalability issues when simulating a Tier-2 site and accuracy issues regarding network load \cite{1742-6596-331-7-072038}.
